---
title: "Code for Azimuth annotation"
output: html_notebook
---

This notebook reports the steps to to annotate the cells for Dataset A and B according to the Azimuth annotation for the article Donada et al.
Please refer to the article for more details.

## Libraries and sources
Load libraries

```{r}
library(tidyverse)
library(Seurat)
library(matrixStats)
library(Azimuth)
```

Set working directory and load routines 
You can set your own path where the file is located, for example:  
setwd("C:\\Users\\UserName\\Documents\\Transcriptomics")

```{r}
script_path <- rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(script_path))
source("repository_functions.R")
```



## Parameters
Select parameters to run Azimuth with a small dataset
```{r}
mapping_score_k <- 7
k_weight <- 7
k_param <- 7 
```

## Load datasets
DatasetA 
```{r}
output_A<-summary_load_v2("../Datasets/dataset1_raw_matrix.csv", 
                          "../Datasets/dataset1_metadata_input.csv")
datasetA<-output_A$obj
datasetA_metadata<-output_A$nm_cells
```

DatasetB 
```{r}
output_B<-summary_load_v2("../Datasets/dataset3_raw_matrix.csv", 
                          "../Datasets/dataset3_metadata_input.csv")
datasetB<-output_B$obj
datasetB_metadata<-output_B$nm_cells
```

Create Seurat Objects
```{r}
expA.realdata <-seurat_object_wt_metadata(datasetA, datasetA_metadata)
expB.realdata <-seurat_object_wt_metadata(datasetB, datasetB_metadata)
```

# Run azimuth
```{r}
expA.azimuth <- expA.realdata
expB.azimuth <- expB.realdata
expA.azimuth <- run_azimuth(expA.azimuth, mapping_score_k, k_weight, k_param)
expB.azimuth <- run_azimuth(expB.azimuth, mapping_score_k, k_weight, k_param)
```
Merge  Azimuth annotation for dataset A and dataset B 
```{r}
merged.ann <-
  rbind(expA.azimuth@meta.data %>% select("predicted.celltype.l2"),
    expB.azimuth@meta.data %>% select("predicted.celltype.l2"))
colnames(merged.ann) <- "separated.azimuth"
```

# Save the new metadata
```{r}
metadataA_complete <- metadata_citeseq_azimuth("../Datasets/dataset1_metadata_input.csv","../Output/dataset1_metadata_step1.csv", expA.azimuth, merged.ann)
metadataB_complete <- metadata_citeseq_azimuth("../Datasets/dataset3_metadata_input.csv", "../Output/dataset3_metadata_step1.csv",expB.azimuth, merged.ann)
```