---
title: "Cosine distance computation"
output: html_notebook
---
This notebook reports the computation of cosine distance and the computation of p-value.
The MIIC network is obtained with the distances computed with the raw counts.
The figures in the article are computed with the normalized counts.

## Libraries and sources
Load libraries
```{r}
library(Seurat)
library(matrixStats)
library(tidyverse)
library(plotly)
library(pals)
library(rstatix)
```

Set working directory and load routines 
You can set your own path where the file is located, for example:  
setwd("C:\\Users\\UserName\\Documents\\Transcriptomics")

```{r}
script_path <- rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(script_path))
source("repository_functions.R")
```


## Parameters
```{r}
which_slot <- "data" # or "counts"
m_seed <- "123" #m_seed = 0
max_y <- ifelse(which_slot == "data", 13, 10)
max_x <- 6

```

## Load datasets
Dataset A  
```{r}
output_A<-summary_load_v2("../Datasets/dataset1_raw_matrix.csv", "../Output/dataset1_metadata_step2.csv")
datasetA<-output_A$obj
datasetA_metadata<-output_A$nm_cells
```

Dataset B
```{r}
output_B<-summary_load_v2("../Datasets/dataset3_raw_matrix.csv", "../Output/dataset3_metadata_step2.csv")
datasetB<-output_B$obj
datasetB_metadata<-output_B$nm_cells
```

Create Seurat objects
```{r}
expA.realdata <-seurat_object_wt_metadata(datasetA, datasetA_metadata)
expB.realdata <-seurat_object_wt_metadata(datasetB, datasetB_metadata)
# Create merged Seurat object
stem.combined <- merge(expA.realdata, y = expB.realdata, add.cell.ids = c("exp1", "exp3"), project = "StemCells")
stem.combined$exp <- sub("_.*", "", rownames(stem.combined@meta.data))
stem.combined@meta.data$family[is.na(stem.combined@meta.data$family)]= 'missing_family'
```

## Compute distances
```{r}
# normalize if needed
if (which_slot=='data'){
  stem.combined <- NormalizeData(stem.combined)}
# find unique families
unique_fam <- unique(stem.combined$family)
# find cell types for each family
df_types <- find_celltype_family(unique_fam, stem.combined)
# find combinations
my_comb <- find_comb(unique_fam, stem.combined)
md_cosine <- c()
for (idx_c in 1:length(my_comb)){
  md_cosine[[names(my_comb)[idx_c]]] <- compute_md_cosine(my_comb[[idx_c]], which_slot)
}
md_df_cosine <- data.frame(t(data.frame(md_cosine)))
colnames(md_df_cosine) <- "distance"
md_df_cosine$family <-rownames(md_df_cosine)
# count number of sequenced cells per family
cts <- dplyr::count(stem.combined@meta.data, family)
md_df_cosine <- (merge(md_df_cosine, cts,
by='family') %>% select(c("distance", "family", "n")))
md_df_cosine <- merge(md_df_cosine, df_types, by='family')
```


## Save files for MIIC analysis
```{r}
if (which_slot=='count'){
add_distance_metadata("../Output/dataset1_metadata_step2.csv", "../Output/dataset1_metadata_final.csv", md_df_cosine)
add_distance_metadata("../Output/dataset3_metadata_step2.csv",  "../Output/dataset3_metadata_final.csv", md_df_cosine)}
```

## Plot cosine distance for real families
```{r}
colors_family <- polychrome(30)
#name_family <- sort(union(expA.realdata$family, expB.realdata$family))
md_df_cosine[md_df_cosine$family=="missing_family","family"] <- "IIA9"
labels <- sort(paste(md_df_cosine$family, " (", md_df_cosine$n, ")", sep=""))

plot1 <- ggplot(md_df_cosine, aes(x=distance, fill=family, label=n, text=cell.fates)) + 
geom_histogram(binwidth=0.05, color='white')+ 
labs(title='Cosine distance: real families', fill='Family')+
scale_x_continuous(limits=c(0,max_x)/10,breaks=seq(0,max_x)/10)+
scale_y_continuous(limits=c(0,max_y), breaks=seq(0,max_y)) + scale_fill_manual(values=unname(colors_family), labels=labels)

plot1
ggplotly(plot1)
```


## Obtain random families
### 10 repetitions: not keeping the cell type and keeping it
```{r}
random_mixed_all_10it <- obtain_random_families(10, unique_fam, stem.combined, F, which_slot, m_seed)
random_type_all_10it <- obtain_random_families(10, unique_fam, stem.combined, T, which_slot, m_seed)
```

### 100 repetitions: not keeping the cell type and keeping it
```{r}
random_mixed_all_100it <- obtain_random_families(100, unique_fam, stem.combined, F, which_slot, m_seed)
random_type_all_100it <- obtain_random_families(100, unique_fam, stem.combined, T, which_slot, m_seed)
```

### 1000 repetitions: not keeping the cell type and keeping it
``` {r}
random_mixed_all_1000it <- obtain_random_families(1000, unique_fam, stem.combined, F, which_slot, m_seed)
random_type_all_1000it <- obtain_random_families(1000, unique_fam, stem.combined, T, which_slot, m_seed)
```

### 10000 repetitions
``` {r}
random_mixed_all_10000it <- obtain_random_families(10000, unique_fam, stem.combined, F, which_slot, m_seed)
random_type_all_10000it <- obtain_random_families(10000, unique_fam, stem.combined,  T, which_slot, m_seed)
```

## Obtain plots for random families  
###Theme for saving figures
``` {r}
default_theme <- theme(plot.title = element_text(hjust =0.5, size=26), axis.line = element_line(size = 2),axis.text = element_text(size = 24),axis.title = element_text(size = 26),legend.text = element_text(size = 22),
legend.title = element_text(size = 24))
``` 

### 10 repetitions
``` {r}
output_10it <- obtain_plots_iteration(random_mixed_all_10it, random_type_all_10it, 0.1, max_x, max_y)

output_10it$plot_mixed_wtrep
output_10it$plot_type_wtrep
```

### 100 repetitions
``` {r}
output_100it <- obtain_plots_iteration(random_mixed_all_100it, random_type_all_100it, 0.01, max_x, max_y)

output_100it$plot_mixed_wtrep
output_100it$plot_type_wtrep

plot2 <- plot1 + theme_classic() + default_theme 
plot_mixed_wtrep <- output_100it$plot_mixed_wtrep + theme_classic() + default_theme
plot_type_wtrep <- output_100it$plot_type_wtrep + theme_classic() + default_theme

```
Save image
``` {r}
ggsave(plot2, 
filename ="FiguresJan/Fig_cosinedistances_realfamilies.svg", device = "svg", height = 6, width = 8.5, units = "in")
ggsave(plot_mixed_wtrep, 
filename ="FiguresJan/Fig_cosinedistances_normalization_random_wofate.svg", device = "svg", height = 6, width = 5, units = "in")
ggsave(plot_type_wtrep, 
filename ="FiguresJan/Fig_cosinedistances_normalization_random_wtfate.svg", device = "svg", height = 6, width = 5, units = "in")
```

### 1000 repetitions
``` {r}
output_1000it <- obtain_plots_iteration(random_mixed_all_1000it, random_type_all_1000it, 0.001, max_x, max_y)

output_1000it$plot_mixed_wtrep
output_1000it$plot_type_wtrep
```

###10.000 repetitions
``` {r}
output_10000it <- obtain_plots_iteration(random_mixed_all_10000it, random_type_all_10000it, 0.0001, max_x, max_y)

output_10000it$plot_mixed_wtrep
output_10000it$plot_type_wtrep
```

# Differences across distributions based on iterations
``` {r}
plot_mixed_diff <- output_10it$plot_mixed_wtrep|output_100it$plot_mixed_wtrep|output_1000it$plot_mixed_wtrep|output_1000it$plot_mixed_wtrep

plot_type_diff <- output_10it$plot_type_wtrep|output_100it$plot_type_wtrep|output_1000it$plot_type_wtrep|output_1000it$plot_type_wtrep

plot_mixed_diff
plot_type_diff

```

# Side by side
``` {r}
h <- plot_distibutions_side_by_side(md_df_cosine, output_100it, n_reps=100, value_weight=0.01, max_x, max_y) 

h_save <- h + theme_classic() +  
scale_fill_manual(values=c("#FFB8B1","#B5EAB6","#DFCCF1"), 
name='Families', labels=c('Real', 'Random', 'Random, same fate')) + default_theme
``` 

Save image
``` {r}
ggsave(h_save, 
filename ="../Figures/Fig_cosinedistances_normalization_sidebyside.svg", device = "svg", height = 6, width = 15, units = "in")
```
