---
title: "Code for MIIC feature selection"
output: html_notebook
---
This notebook reports the feature selection for the MIIC networks presented in the article Donada et al. 
Please refer to the article for more details.

## Libraries and sources
Load libraries

```{r}
library(Seurat)
library(tidyverse)
library(matrixStats)
library(miic) #Version used: 1.9.0
```

Set working directory and load routines
You can set your own path where the file is located, for example:  
setwd("C:\\Users\\UserName\\Documents\\Transcriptomics")
```{r}
script_path <- rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(script_path))
source("repository_functions.R")
```

## Parameters
if flag_avg=TRUE, then we compute the mean of the MT-, RPL, RPS, MRPLS, MRPS genes.
In the MIIC networks obtained for the article, flag_avg <-0

```{r}
flag_avg <- 0
# variables that are regulators
vec_TF <-c("family", 
"T1", "avg_dT2", "avg_dT3", 
#"max_var", "min_var", 
"predicted.celltype",
"Colony_size", 
#"homogeneity", "n_celltypes", 
"exp","distance")
# thresholds for filtering
th1 <- 0.06
th2 <- th1*2.4
```

## Load datasets
DatasetA
```{r}
output_A<-summary_load_v2("../Datasets/dataset1_raw_matrix.csv", "../Output/dataset1_metadata_final.csv")
datasetA<-output_A$obj
datasetA_metadata<-output_A$nm_cells
```   

DatasetB
```{r}
output_B<-summary_load_v2("../Datasets/dataset3_raw_matrix.csv", "../Output/dataset3_metadata_final.csv")
datasetB<-output_B$obj
datasetB_metadata<-output_B$nm_cells
```

## Create Seurat objects
```{r}
expA.realdata <-seurat_object_wt_metadata(datasetA, datasetA_metadata)
expB.realdata <-seurat_object_wt_metadata(datasetB, datasetB_metadata)
```

Merge data
```{r}
stem.combined <- merge(expA.realdata, y = expB.realdata, add.cell.ids = c("exp1", "exp3"), project = "StemCells")
stem.combined$exp <- sub("_.*", "", rownames(stem.combined@meta.data))
```

## Add the variables "homogeneity" and "number of cells" 
which encode the fate similarity inside the same family. 
```{r}
#stem.combined$n_cells <- rownames(stem.combined@meta.data)
#new_vars<- compute_homogeneity(stem.combined@meta.data) 
#rownames(new_vars) <- new_vars$n_cells
```

## Prepare input for feature selection
```{r}
#df.merged <- merge(stem.combined@meta.data, 
#               new_vars[, c("homogeneity", "n_celltypes")],
#               by = "row.names")

df.merged <- stem.combined@meta.data
df.merged <- df.merged[,vec_TF]
new_order <-match(row.names(df.merged), colnames(stem.combined))
df.merged <- df.merged[order(new_order),]
datasets <- data.frame(stem.combined@assays$RNA@counts)
#match(row.names(df.merged), colnames(datasets))
df.merged[is.na(df.merged$family),]$family <- "missing_family"
```

## Run feature selection
```{r}
if (flag_avg){
all_groups <- c("^MT-", "^RPL", "^RPS", "^MRPL", "^MRPS")
df_output <- compute_avg_mitoribo(stem.combined, all_groups)
rm(datasets)
datasets <- df_output
# recreate new Seurat object
df_metadata <- stem.combined@meta.data
#rownames(df_metadata) <- gsub(".*_", "", rownames(df_metadata))
stopifnot(identical(colnames(df_output), rownames(df_metadata)))
stem.new <- CreateSeuratObject(counts = df_output, 
row.names = rownames(df_output))
stem.new <- AddMetaData(object = stem.new, metadata = df_metadata)
stem.combined <- stem.new}

all_res <- list()
for (i in 1:length(vec_TF)) {
  one_TF <- vec_TF[i]
  print(one_TF)
  res <-
    seurat_mi_compute(
      mat = datasets,
      value_name = one_TF,
      values = df.merged[,one_TF],
      cell_types = NULL,
      tfs = NULL,
      recompute = F,
      display = T,
      pipeline = F
    )
  all_res[[one_TF]] <- res
}
```

## Write table with info_shifted for each regulator 
```{r}
df <- all_res
df <- do.call(cbind, df)
```

Save file 
```{r}
write.csv(df, "../Output/df_for_fs.csv", row.names=T)
```

## Filtering
```{r}
rm(df)
# Reload dataframe from saved file
df <- read.csv( "../Output/df_for_fs.csv", header=T,row.names="X")

# Delete mitochondrial and ribosomial genes, 
if (!flag_avg){
to_delete <- c("^MT-|^RPL|^RPS|^MRPL|^MRPS")
} else {
 to_delete <- NULL
}

# Filter variables
vars <- filter_info(df,th1, to_delete)
vars2 <- filter_info(df,th2, to_delete)
vars3 <- filter_info(df%>%select(-c("family")),th1, to_delete)
vars4 <- filter_info(df%>%select(-c("family")),th2, to_delete)
```

## Save input
Save input and check equality with original file.
```{r}
# for network with the family node
df_family <- save_input(vec_TF, stem.combined, df.merged, 
vars, vars2, "wtfamily", "MIIC")
equal_check(df_family, datasetA, "exp1")
equal_check(df_family, datasetB, "exp3")
# for network without the family node
df_wo_family <- save_input(vec_TF[-1], stem.combined, df.merged%>% select(-c("family")), vars3, vars4, 'wofamily', "MIIC")
equal_check(df_wo_family, datasetA, "exp1")
equal_check(df_wo_family, datasetB, "exp3")

```
